Debugging of RxJS-based Applications

### 2020-11-02 paper.tex
Rework Walkthrough

\begin{itemize}
	\item Wharton et al. \cite{Wharton_Rieman_Clayton_Polson_1994}
	\item See Appendix~\ref{sec:persona} for full detail on the results of the cognitive walkthrough.
	\item Present setup
\end{itemize}

@@

\appendix
\section{Persona ``Roger Flow''}
\label{sec:persona}

@@

\subsubsection{User}

See Appendix~\ref{sec:persona} ``Persona Roger Flow''

\subsubsection{Environment}
Visual Studio Code is installed on an arbitrary system. Support for TypeScript projects is enabled as well as the proof of concept implementation of our RxJS debugging extension is installed and enabled. A launch configuration to execute the project in the systems default browser is available. Further, an internet browser (e.g. Mozilla Firefox or Google Chrome) is available.

@@

TODO Make Measurable! SMART?

I want to understand the values and life cycle events produced by the flatMap operator in the \texttt{index.ts} file of \texttt{problem-1}\footnote{\url{https://github.com/swissmanu/mse-pa1-experiment/blob/v1.0.2/src/problem-1/index.ts\#L18}} and how they affect the whole application.

\subsubsection{Action Sequence}

\subsection{Step-by-Step Analysis Phase}

@@

\begin{enumerate}
	\item Open \texttt{index.ts} in Visual Studio Code
	      \begin{itemize}
	      	\item Visual Studio Code: Shows contents of \texttt{index.ts} file
	      	\item Success story:
	      	      \begin{itemize}
	      	      	\item This seems OK based on prior experience with Visual Studio code.
	      	      \end{itemize}
	      \end{itemize}

	\item Move cursor the \texttt{flatMap} operator on line 18
	      \begin{itemize}
	      	\item Visual Studio Code: Show code actions icon in front of line 18
	      	\item Success story:
	      	      \begin{itemize}
	      	      	\item This seems OK because the original task clearly states a question regarding this line/piece of source code. Hence, navigating here seems like the natural course of action.
	      	      \end{itemize}
	      \end{itemize}

	\item Open code actions menu
	      \begin{itemize}
	      	\item Visual Studio Code: Show available code actions
	      	\item Failure story:
	      	      \begin{itemize}
	      	      	\item Criterion: Will the user know that the correct action is available?
	      	      	      \begin{itemize}
	      	      	      	\item The user might know code actions for providing options to refactor a piece of code or quick fixes for code linting problems. It is questionable if they will expect functionality to inspect parts of a data flow graph here.
	      	      	      \end{itemize}
	      	      \end{itemize}
	      \end{itemize}

	\item Select "Probe Observable..." code actions
	      \begin{itemize}
	      	\item Visual Studio Code: Add \texttt{flatMap} operator on line 18 to "Observables" list in debugging view
	      	\item Failure story:
	      	      \begin{itemize}
	      	      	\item Criterion: If the correct action is taken, will the user see that things are going ok?
	      	      	      \begin{itemize}
	      	      	      	\item The "Observables" list is part of the debugging view of Visual Studio Code. The user will not get any feedback that their action "Probe Observable..." was successful without changing the view manually to debugging and expanding the "Observables" panel in the lower left.
	      	      	      \end{itemize}
	      	      \end{itemize}
	      \end{itemize}

	\item Open "Observable Probe Monitor" view using command palette
	      \begin{itemize}
	      	\item Visual Studio Code: Show empty "Observable Probe Monitor" view
	      	\item Failure story:
	      	      \begin{itemize}
	      	      	\item Criterion: Will the user know that the correct action is available?
	      	      	      \begin{itemize}
	      	      	      	\item A user might not be aware that the "Observable Probe Monitor" view is hidden as a command in the command palette. Hence, they might feel lost after adding the observable probe in the previous step.
	      	      	      \end{itemize}
	      	      	\item Criterion: If the correct action is taken, will the user see that things are going ok?
	      	      	      \begin{itemize}
	      	      	      	\item The user might get confused by the "Observable Probe Monitor" being blank by default.
	      	      	      \end{itemize}
	      	      \end{itemize}
	      \end{itemize}

	\item Execute "Problem 1" launch configuration
	      \begin{itemize}
	      	\item Visual Studio Code: Open default browser showing "Problem 1"
	      	\item Default Browser: Show "Problem 1"
	      	\item Success story:
	      	      \begin{itemize}
	      	      	\item This seems OK according the past experiences with Visual Studio Code's launch configurations.
	      	      \end{itemize}
	      \end{itemize}

	\item Interact with "Problem 1" in the default browser
	      \begin{itemize}
	      	\item Visual Studio Code: "Observable Probe Monitor" traces values and life cycle events produced by the \texttt{flatMap} operator
	      	\item Failure story:
	      	      \begin{itemize}
	      	      	\item Criterion: Will the user know that the correct action will achieve the desired effect?
	      	      	      \begin{itemize}
	      	      	      	\item The user might not be aware that they are expected to interact with "Problem 1" in the default browser in order to get live feedback in the "Observable Probe Monitor" of emitted values and life cycle events.
	      	      	      \end{itemize}
	      	      	\item Criterion: If the correct action is taken, will the user see that things are going ok?
	      	      	      \begin{itemize}
	      	      	      	\item The default browser might overlay Visual Studio Code and the "Observable Probe Monitor" view. This is why the user might miss the live trace of values and life cycle events logged to the "Observable Probe Monitor" view.
	      	      	      \end{itemize}
	      	      \end{itemize}
	      \end{itemize}

	\item Interpret the live trace of emitted values and life cycle events in the "Observable Probe Monitor" view
	      \begin{itemize}
	      	\item Visual Studio Code: Provide detail information to a traced item
	      	\item Success story:
	      	      \begin{itemize}
	      	      	\item This seems OK based on the original task where the user wants to understand runtime behavior of the `flatMap` operator.
	      	      \end{itemize}
	      \end{itemize}
\end{enumerate}

>>>>

\begin{itemize}
	\item Wharton et al. \cite{Wharton_Rieman_Clayton_Polson_1994}
	\item See Appendix~\ref{sec:cogitive-walkthrough-appendix} and for full detail on the results of the cognitive walkthroughs.
	\item Present setup
\end{itemize}

@@

\appendix

\section{Persona ``Roger Flow''}
\label{sec:persona}

@@

This cognitive walkthrough is based on the first problem given to subjects during the observational study of our previous work ``Debugging RxJS-based Applications''\cite{TODO}.

\subsubsection{User}
See Appendix~\ref{sec:persona} ``Persona Roger Flow''

@@

After I started the ``Problem~1'' application and inspected its UI, I was able to observe multiple, unexpected updates rendered in quick succession after I clicked the reset button. Based on this evidence, I formulate my first debugging hypothesis: I suspect that the \texttt{flatMap} operator on Line~18 in the file \texttt{index.ts}\footnote{\url{https://github.com/swissmanu/mse-pa1-experiment/blob/v1.0.2/src/problem-1/index.ts\#L18}} does create multiple observables, which do not get unsubscribed when the reset button is clicked. This results in the observed behavior eventually. To proof my hypothesis, I want to inspect the life cycle events of the created observables more closely.

\subsubsection{Environment}

Visual Studio Code with enabled TypeScript support is installed. The prototype of our RxJS debugging extension is installed as well. The source code of Problem~1 from our previous experiment is available. Further, an internet browser (e.g. Mozilla Firefox or Google Chrome) is present.

\subsection{Walkthrough}

@@

\subsubsection{Open File}
Open \texttt{index.ts} in Visual Studio Code.

\begin{itemize}
	\item Visual Studio Code: Shows contents of \texttt{index.ts} file.
	\item Success story:
	      \begin{itemize}
	      	\item We can expect the user to open \texttt{index.ts} since he already suspects a problem within this file as stated in the original task.
	      \end{itemize}
\end{itemize}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth]{walkthrough-screenshots/step1.png}
	\Description{}
	\caption{Visual Studio Code after opening the \texttt{index.ts} file.}
	\label{fig:walkthrough-screesnhot-step-1}
\end{figure}

\subsubsection{Navigate to Operator}
Move cursor the \texttt{flatMap} operator on Line~18.

\begin{itemize}
	\item Visual Studio Code: Shows code actions icon in front of Line~18.
	\item Success story:
	      \begin{itemize}
	      	\item The original task clearly describes the hypothesis regarding this line/piece of source code. Hence, navigating here seems the natural course of action for the user.
	      \end{itemize}
\end{itemize}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth]{walkthrough-screenshots/step2.png}
	\Description{}
	\caption{Visual Studio Code after navigating cursor to the \texttt{flatMap} operator on Line~18.}
	\label{fig:walkthrough-screesnhot-step-2}
\end{figure}

\subsubsection{Open Code Actions}
Open the code actions menu by clicking the yellow light bulb icon.

\begin{itemize}
	\item Visual Studio Code: Shows available code actions.
	\item Failure story:
	      \begin{itemize}
	      	\item Will the user know that the correct action is available?
	      	      \begin{itemize}
	      	      	\item The user might know code actions for providing options to refactor a piece of code or quick fixes for code linting problems. It is questionable if he will expect functionality to inspect parts of a data flow graph here.
	      	      \end{itemize}
	      \end{itemize}
\end{itemize}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth]{walkthrough-screenshots/step3and4.png}
	\Description{}
	\caption{Visual Studio Code indicating available code actions on Line~18 using a yellow light bulb icon.}
	\label{fig:walkthrough-screesnhot-step-3}
\end{figure}

\subsubsection{Create Probe for Operator}
Select ``Probe Observable...'' code action from the related menu.

\begin{itemize}
	\item Visual Studio Code: Adds \texttt{flatMap} operator on Line~18 to "Observables" list in debugging view.
	\item Failure story:
	      \begin{itemize}
	      	\item If the correct action is taken, will the user see that things are going ok?
	      	      \begin{itemize}
	      	      	\item The ``Observables'' list is part of the debugging view of Visual Studio Code. The user will not get any feedback that his action ``Probe Observable...'' was successful without changing the view manually to debugging and expanding the ``Observables'' panel in the lower left.
	      	      \end{itemize}
	      \end{itemize}
\end{itemize}

\subsubsection{Open Observable Probe Monitor}
Open the ``Observable Probe Monitor'' view using command palette.

\begin{itemize}
	\item Visual Studio Code: Shows empty ``Observable Probe Monitor'' view
	\item Failure story:
	      \begin{itemize}
	      	\item Will the user know that the correct action is available?
	      	      \begin{itemize}
	      	      	\item The user might not be aware that the ``Observable Probe Monitor'' view is hidden within the command palette. Hence, they might feel lost after adding the observable probe in the previous step.
	      	      \end{itemize}
	      	\item If the correct action is taken, will the user see that things are going ok?
	      	      \begin{itemize}
	      	      	\item The user might get confused by the ``Observable Probe Monitor'' being blank by default.
	      	      \end{itemize}
	      \end{itemize}
\end{itemize}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth]{walkthrough-screenshots/step5-1.png}
	\Description{}
	\caption{Visual Studio Codes command palette menu showing the ``Observable Probe Monitor'' command.}
	\label{fig:walkthrough-screesnhot-step-5-1}
\end{figure}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth]{walkthrough-screenshots/step5-2.png}
	\Description{}
	\caption{Visual Studio Code showing the empty Observable Probe Monitor on the right pane.}
	\label{fig:walkthrough-screesnhot-step-5-2}
\end{figure}

\subsubsection{Launch Application}
Execute ``Problem~1'' launch configuration

\begin{itemize}
	\item Visual Studio Code: Opens default browser showing ``Problem~1''
	\item Default Browser: Shows ``Problem~1'' UI
	\item Success story:
	      \begin{itemize}
	      	\item The users previous experience with Visual Studio Code launch configuration allows assuming this the natural course of action in order to prepare himself for further inspection of the application.
	      \end{itemize}
\end{itemize}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth]{walkthrough-screenshots/step6.png}
	\Description{}
	\caption{Visual Studi Code showing the debugging view after launching ``Problem~1''.}
	\label{fig:walkthrough-screesnhot-step-6}
\end{figure}

\subsubsection{Interact with Application}
Interact with ``Problem~1'' in the default browser.

\begin{itemize}
	\item Visual Studio Code: ``Observable Probe Monitor'' provides live telemetry information about values and life cycle events produced by the \texttt{flatMap} operator.
	\item Failure story:
	      \begin{itemize}
	      	\item Will the user know that the correct action will achieve the desired effect?
	      	      \begin{itemize}
	      	      	\item The user might not be aware that he is expected to interact with ``Problem~1'' in the default browser in order to get live feedback in the ``Observable Probe Monitor''.
	      	      \end{itemize}
	      	\item If the correct action is taken, will the user see that things are going ok?
	      	      \begin{itemize}
	      	      	\item The default browser might overlay Visual Studio Code and the ``Observable Probe Monitor'' view. This is why the user might miss the live trace of values and life cycle events displayed in the ``Observable Probe Monitor''.
	      	      \end{itemize}
	      \end{itemize}
\end{itemize}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth]{walkthrough-screenshots/step7.png}
	\Description{}
	\caption{Google Chrome displaying the user interface of ``Problem~1'' ready to receive interactions.}
	\label{fig:walkthrough-screesnhot-step-7}
\end{figure}

\subsubsection{Interpret Runtime Behavior}
Interpret the live trace of emitted values and life cycle events in the ``Observable Probe Monitor'' view

\begin{itemize}
	\item Visual Studio Code: Provides detail information to a traced item
	\item Success story:
	      \begin{itemize}
	      	\item The original task states that the user is interested in more close information regarding the \texttt{flatMap} operator. Since the ``Observable Probe Monitor'' provide such information in real-time, we can expect the user to use this information accordingly.
	      \end{itemize}
\end{itemize}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\columnwidth]{walkthrough-screenshots/step8.png}
	\Description{}
	\caption{Visual Studio Code showing live telemetry in the ``Observable Probe Monitor''.}
	\label{fig:walkthrough-screesnhot-step-8}
\end{figure}
